---
title: "Survey Simulation Example"
author: "Eura Nama"
output:
  # bookdown::word_document2:
  #   fig_caption: yes
  #   number_sections: false
  # fontsize: 12pt
  # sansfont: Liberation Sans
  # mainfont: Liberation Sans
  # classoption: twocolumn
  # language: english
  #bookdown::html_document2: default
  bookdown::pdf_document2:
      keep_tex: yes
      number_sections: false
      toc: no
      # includes:
      # in_header: "styling.sty"
  language: english
  classoption: twocolumn
header-includes: 
  - \usepackage{tikz} \usepackage{pdflscape} \usepackage{float}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
---



<!-- Here we are setting the parameters for our modelling -->

```{r survey-parameters, echo=F}
options(scipen = 999) # Just forces numbers to not be in scientific notiation
# We can set the number of stations...
n.stations <- 20
tot.biomass <- 100000 # Could be any units, let's call it tonnes, and because of how I have this set up below this number isn't exact
catchability <- 0.3 # What proportion of the population does survey tend to catch, going to fix this at 0.3 for simplicity
area.swept <-  10000    # What is the area swept of each tow, we'll assume it is identical, this would be a 2 km run with 5 meter wide gear.  
                                  # somewhere in the 1000's of square meters tends to make sense here for trawl kinds of gear.  Also need to convert to km²
n.sims <- 4
bm.dist <- c("Random","NAFO","Depth") 
# Which Biomass distribution do you want [1] will be Random Survey, 2 will be NAFO survey, and 3 will be "Depth
bm.dist <- bm.dist[2]


```

```{r setup, include=FALSE,echo=F, message=F,warning=F,cache=F}

# First up we will check your r packages and install anything you need for this for you.
req.packages <- c("tidyverse","lubridate","plotly","sf","sp","data.table","units","cowplot","knitr",'concaveman','ggthemes',"nngeo","marmap","RandomFields",
                	"ggplot2","stars","tmaptools","rnaturalearth","rnaturalearthdata","raster","rgdal","RStoolbox","pals","ggnewscale","ggspatial",'devtools','rlist')
# If you don't have the packages install them + give a heads up that you are
new.packages <- req.packages[!(req.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)>0) 
{
  cat(paste0("Heads up, I have to install these packages for this to work:", new.packages ))
  #wanna.install <- readline(prompt = "If you want to install these package(s) enter 'y': ")
  #if(tolower(wanna.install) == 'y') 
  install.packages(new.packages,repos = "http://cran.us.r-project.org") #else { stop("You didn't want to install the packages so this script does not work.")}
}

# You also need to install this github repo package if you do not have it.
hi.res <- any(installed.packages()[,"Package"] %in% "rnaturalearthhires")
if(hi.res == F) devtools::install_github("https://github.com/ropensci/rnaturalearthhires/")
  

# The libraries that we will use directly in this code
library(knitr)
library(ggplot2)
library(gridExtra)
library(reshape2)
library(scales)
library(cowplot)
require(tidyverse)
require(dplyr)
require(tidyr)
library(RandomFields)
library(sf)
library(marmap)
library(stars)
library(raster)

knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
# Set the Workding directory.
direct.proj <- "D:/Github/Fisheries_module/"

#Some Custom Functions I'll need
funs <- c("https://raw.githubusercontent.com/Mar-Scal/Assessment_fns/master/Maps/pectinid_projector_sf.R",
          "https://raw.githubusercontent.com/Mar-scal/Assessment_fns/master/Maps/centre_of_gravity.R",
          "https://raw.githubusercontent.com/Mar-scal/Assessment_fns/master/Maps/add_alpha_function.R",
          "https://raw.githubusercontent.com/Mar-scal/Assessment_fns/master/Maps/combo_shp.R",
          "https://raw.githubusercontent.com/Mar-scal/Assessment_fns/master/Maps/convert_coords.R")
# Now run through a quick loop to load each one, just be sure that your working directory is read/write!
for(fun in funs) 
{
  download.file(fun,destfile = basename(fun))
  source(paste0(getwd(),"/",basename(fun)))
  file.remove(paste0(getwd(),"/",basename(fun)))
}

# A couple custom functions I may or may not use
factor.2.number <- function(x) {as.numeric(levels(x))[x]} # My friend factor.2.number
# Function in case you need it for transforming propotion data to not have 0's and 1's.  
beta.transform <- function(dat,s=0.5)  (dat*(length(dat)-1) + s) / length(dat)
# Cute little function to divide split something up by a weighting factor.
nsplit = function(X,n)
{
  p = X/sum(X)
  diff(round(n*cumsum(c(0,p))))
}
```


```{r data-setup,echo=F, cache=T,include=F,message=F}
# Now I want to make a bounding box that outlines our study area
survey.domain <- st_as_sf(data.frame(X = c( 465000, 750000, 750000, 650000, 600000, 400000),
                                    Y = c(4450000,4450000,4710000,4900000,4860000,4800000),ID=1),coords = c("X","Y"),crs= 32619)
# Transform it to proper units
survey.domain <- st_transform(survey.domain,crs = 4326)
# And convert it from some points into a polygon object
survey.domain <- st_cast(st_combine(survey.domain),"POLYGON")

# Now we can carve up our survey domain into a bunch of grids, size is I believe 0.1 degrees, so 6 minutes each I think (10 per degree)
survey.grid <- st_make_grid(survey.domain,cellsize = 0.1)
# Then we clip that to the survey domain and we have some nice cells we can toss biomass into kinda almost anyway we'd like.
survey.grid <- st_intersection(survey.grid,survey.domain)
# Make this a nice sf object for later.
survey.grid <- st_sf(survey.grid)


# Lets grab the NAFO subareas that are going to form our survey domain, I'll need to tweak a few of them to cut off to the south of GB
# Figure out where your tempfiles are stored
temp <- tempfile()
# Download this to the temp directory you created above
download.file("https://raw.githubusercontent.com/Mar-scal/GIS_layers/master/NAFO/Subareas.zip", temp, quiet=T)
# Figure out what this file was saved as
temp2 <- tempfile()
# Unzip it
unzip(zipfile=temp, exdir=temp2)
# This pulls in all the layers from the above location
nafo <- combo.shp(temp2,make.sf=T, quiet=T)
nafo <- st_make_valid(nafo)
# Now clip the nafo strata into our made up survey domain
nafo.strata <- st_intersection(nafo,survey.domain)
nafo.strata
# Two of the strata kinda suck, id 125, 135, 141, and 142 are going to go as they aren't unique subareas.
nafo.strata <- nafo.strata %>% dplyr::filter(!id %in% c(125,135,141,142))
# Calculate the total area and the area by strata
nafo.strata$area <- st_area(nafo.strata) %>% units::set_units("km^2") %>% as.numeric()
tot.area <- sum(nafo.strata$area) # This gets the same number as using st_area(survey.domain) so this is the total area of the survey domain
# Now we calculate the proportion of the total area that each strata covers
nafo.strata$p.area <-nafo.strata$area/tot.area

#Lots of other things you can stratify on as well, a classic is to stratify by depth, we can do that by using the bathymetry that we have in the above plots
# which is completely free to download  

bathy.org <- getNOAA.bathy(lon1 = -72 ,lon2 = 65,lat1 = 40,lat2=44.5,resolution =1)
bathy <- marmap::as.raster(bathy.org)
b.box <- st_bbox(survey.domain)
# Now clip this to the bounding area, note that the bathy is basically a EPSG:4326 so we need to crop it accordingly and transform our b.box to this...
bathy.s <- crop(bathy,b.box)
bathy.st <- st_as_stars(bathy.s)
bathy.poly <- st_as_sf(bathy.st)
# Make anything > 0 = 0, and anything < 250 = 250.  Our species won't go below 250 meters depth nor is it an amphibian
bathy.poly$layer[bathy.poly$layer > 0] <- 0
bathy.poly$layer[bathy.poly$layer < -250] <- -250
bathy.poly$strata <- NA
# Now we can simplify the object further so there isn't a polygon for every 1 meter difference, here we lump things into some nicer bins
# Going to make depth contours every 20 meters.  Sure there is a nicer way to do this, but I don't know it...

bathy.poly$strata[bathy.poly$layer >= -25]                          <- "< 25 meters" # Make everything great that 25 meter depth a strata
bathy.poly$strata[bathy.poly$layer >= -50 & bathy.poly$layer < -25] <- "25 to 50 meters"
bathy.poly$strata[bathy.poly$layer >= -75 & bathy.poly$layer < -50] <- "50 to 75 meters"
bathy.poly$strata[bathy.poly$layer >=-100 & bathy.poly$layer < -75] <- "75 to 100 meters"
bathy.poly$strata[bathy.poly$layer >=-125 & bathy.poly$layer < -100] <- "100 to 125 meters"
bathy.poly$strata[bathy.poly$layer >=-150 & bathy.poly$layer < -125] <- "125 to 150 meters"
bathy.poly$strata[bathy.poly$layer >=-175 & bathy.poly$layer < -150] <- "150 to 175 meters"
bathy.poly$strata[bathy.poly$layer >=-200 & bathy.poly$layer < -175] <- "175 to 200 meters"
bathy.poly$strata[bathy.poly$layer >=-225 & bathy.poly$layer < -200] <- "200 to 225 meters"
bathy.poly$strata[bathy.poly$layer >=-250 & bathy.poly$layer < -225] <- "225 to 250 meters"
bathy.poly$strata[bathy.poly$layer <=-250]                           <- "> 250 meters"

# Now make a multi with each 25 meter depth strata a multipolygon object
depth.strata <- aggregate(bathy.poly, list(bathy.poly$strata), function(x) x[1])
# and now clip this to our survey domain (again, last time we used a raster bounding box to trim down, now we really clip it properly)
depth.strata <- st_intersection(depth.strata,survey.domain)
# And calculate the area of each bathy strata
depth.strata$area <- st_area(depth.strata) %>% units::set_units("km^2") %>% as.numeric()
tot.area.bathy <- sum(depth.strata$area) # It is within 3 km^2 of the survey domain, I'll take that!
depth.strata$p.area <- depth.strata$area /tot.area.bathy


```

<!-- This code makes and saves the base plots, but does not display them -->

```{r base-plts,echo=F, warning =F, message=F,include=F,cache=T}
# This code makes and saves the base plots, but does not display them
base.plt <- pecjector(area="GOM",repo = 'github', plot=F,
                      add_layer = list(land ='grey',eez = 'eez',nafo = 'sub',scale.bar = 'tl',bathy = c(10,'s',250)),
                      add_custom = list(obj = survey.domain,fill=NA,size=2,color='darkgrey'))
# We'll also make this plot which we'll want for later, but won't plot now.
base.no.bathy <- pecjector(area="GOM",repo = 'github', plot=F,
                       add_layer = list(land ='grey',nafo = 'sub',scale.bar = 'tl'),
                       add_custom = list(obj = survey.domain,fill=NA,size=2,color='darkgrey'))

```



# Survey Parameters

Just so you can keep track in the document, we return the survey parameters that you set for this simulation.

```{r paras, echo=F}

# Now some derived parameters based on the area we are studying and the inputs provided at the top.
m2.to.km2 <- 1e-6 # convert from m2 to km2
area.swept <- area.swept*m2.to.km2 # Convert area swept by each tow into km²
mn.bm.dens <- tot.biomass/tot.area# The mean biomass density in the region
tot.towable.area <- tot.area / area.swept # This is the total possible number of tows one could do in the entire area.
expected.bm.per.tow <- tot.biomass/tot.area*area.swept*1000*catchability # This is what we should expect to be the biomass per tow in KG given the above inputs.
expect.survey.bm <- tot.biomass*catchability # This is what the biomass from the survey should be given the catchability.


# Let's get all the parameters formatted nicely for a table so we can see them
parameters <- data.frame(Parameter = c("Number of Tows ", "Total Biomass", 
                                       "Catchability","Area swept by a tow",
                                       "Number of Simulations","Biomass distribution"),
                         Value     = c(     n.stations  ,    paste(tot.biomass, 'tonnes')  , 
                                            catchability , paste(area.swept/m2.to.km2,"m²"),
                                            n.sims ,              bm.dist))

knitr::kable(parameters,booktabs=T, caption = "A Table of your input values for the current run of your simulation")

```


```{r survey-simulate, include=FALSE,cache=F}

# Before we loop, we can get the NAFO and Depth areas and number of stations sorted out
nafo.strata$stations <- nsplit(nafo.strata$p.area, n.stations) 
# Now calculate the area swept in each of the areas, use a few metrics here
nafo.strata <- nafo.strata %>% dplyr::mutate(area.surveyed = stations * area.swept,
                                             towable.area = area/area.swept)

#And now we can get ready to stratify by depth
depth.strata$stations <-  nsplit(depth.strata$p.area, n.stations) 
# Now calculate the area swept in each of the areas, use a few metrics here
depth.strata <- depth.strata %>% dplyr::mutate(area.surveyed = stations * area.swept,
                                             towable.area = area/area.swept)

rand.samp <- NULL
nafo.samp <- NULL
depth.samp <- NULL
for(i in 1:n.sims)
{
# Now we can set up location of our samples based on the strata
rand.samp[[i]] <- survey.domain %>% st_sample(size = n.stations) %>% st_sf()

#But we could also have it to be stratified by the NAFO subarea too


# Grab samples, get a new ID field and combine back with the data in the nafo strata object
nafo.samp[[i]] <- st_sample(nafo.strata, size = nafo.strata$stations,type = 'random', exact = T) %>%
                        st_sf('ID' = seq(length(.)), 'geometry' = .) %>%
                        st_intersection(., nafo.strata)


# strat.plt <- base.plt + geom_sf(data = nafo.samp)
# strat.plt


depth.samp[[i]] <- st_sample(depth.strata, size = depth.strata$stations,type = 'random', exact = T) %>%
                        st_sf('ID' = seq(length(.)), 'geometry' = .) %>%
                        st_intersection(., depth.strata)
depth.samp[[i]]$id <- depth.samp$ID
}

```

<!-- We now distribute the biomass across the area according to the bm distribution scheme we set up.  Note that underlying the distribution
     is a Gaussian random field with set characteristics, that would be interesting to play with that to see how that effects things, but that's not for now :-)
-->

```{r biomass-distros, include = F,echo=F,cache=F}
# opts_chunk$set(cache.extra = knitr::rand_seed)
#RFoptions(seed=as.integer(runif(1,-1e6,1e6)))
# So here we make a gmrf which distributes our biomass across the area, this will effectively give us random spatial variability within the random cells.
centroids <- st_centroid(survey.grid)# %>% data.frame()
cent.split <- centroids %>% dplyr::summarise(lat = unlist(map(centroids$survey.grid,1)),
                                      long = unlist(map(centroids$survey.grid,2)))

# Make the bm.res object
bm.res <- NULL
for(i in 1:n.sims)
{
gmrf.pnts <- RFsimulate(model = RMgauss(var=4, scale =2), x=cent.split$lat, y=cent.split$long, grid=F)

gmrf.pnts.sf <- st_as_sf(gmrf.pnts, crs = 4326)
st_crs(gmrf.pnts.sf) <- 4326


if(bm.dist == "Random")
{
# Now we should be able to merge back to the survey grid, I think...
gmrf.bm <- st_join(survey.grid,gmrf.pnts.sf)
# And when we plot it we get what we want... sweet
#ggplot(gmrf.bm) + geom_sf(aes(fill= variable1)) + scale_fill_viridis_b()

# Lets get the size of each grid
gmrf.bm$grid.area <- gmrf.bm %>% st_area() %>% units::set_units("km^2") %>% as.numeric()
# So now we can distribute the biomass based on the gmrf.bm object
gmrf.bm$mn.bm.dens <- mn.bm.dens
# Now we can adjust the biomass density using the gmrf values, not sure this really makes sense, but seems to be fairly reasonable
gmrf.bm$bm.dens <- exp(log(gmrf.bm$mn.bm.dens) + gmrf.bm$variable1)
# So we don't get exactly the same biomass because of how I'm doing the above, but we can corret this back to number we want in a second...
gmrf.bm$biomass <- gmrf.bm$grid.area*gmrf.bm$bm.dens
tmp.biomass <- sum(gmrf.bm$biomass)
# Now we adjust it back to our initial biomass
gmrf.bm$biomass <- gmrf.bm$biomass*tot.biomass/tmp.biomass
gmrf.bm$bm.dens <- gmrf.bm$biomass/gmrf.bm$grid.area
bm.res[[i]] <- gmrf.bm
} #end if(bm.dist == "Random")


# If we are looking at NAFO bimoass distribution do this
if(bm.dist == "NAFO")
{
# So we first build a "NAFO Strata effect". I want this to be similar in size to our GMRF effect above, so somewhere in the -3 to 2 range should be fine
# Here are some effect sizes for this, should be on par with the effect size of the GRMF with the default settings
nafo.strata$effect <- c( 0.25,  #130
                        -2.50,  #131
                        -3.00,  #136
                        -2.75,  #137
                        -2.50,  #138
                        -1.50,  #139
                         0.75,  #140
                         1.00,   #143
                         2.0,   #144
                         1.25,  #145
                         1.00,  #146
                         1.50,  #147
                         1.25)  #148
# Now the nafo strata and our grid 'boxes' aren't going to perfectly overlap, but we can do some gymnastics to have the strata effect due to nafo because
# we did the above.... Start with the points...
nafo.gmrf.pnts <- st_intersection(gmrf.pnts.sf,nafo.strata)
nafo.gmrf.pnts$mn.bm.dens <- mn.bm.dens
nafo.gmrf.pnts$bm.dens <- exp(log(nafo.gmrf.pnts$mn.bm.dens) + nafo.gmrf.pnts$variable1 + nafo.gmrf.pnts$effect)
# Now we can take out the stratification indicators (seems to cause grief with next step)
nafo.gmrf.pnts <- nafo.gmrf.pnts %>% dplyr::select(!c('id','level_0','level_1',"level_2","level_3","ID","area",'stations','area.surveyed','towable.area','p.area'))
# Now put these into the survey grid
gmrf.bm <- st_join(survey.grid,nafo.gmrf.pnts)
#calculate the area of the cells
gmrf.bm$grid.area <- gmrf.bm %>% st_area() %>% units::set_units("km^2") %>% as.numeric()
# And have fun calculating biomasses
gmrf.bm$biomass <- gmrf.bm$grid.area*gmrf.bm$bm.dens
tmp.biomass <- sum(gmrf.bm$biomass)
# Now we adjust it back to our initial biomass
gmrf.bm$biomass <- gmrf.bm$biomass*tot.biomass/tmp.biomass
gmrf.bm$bm.dens <- gmrf.bm$biomass/gmrf.bm$grid.area
bm.res[[i]] <- gmrf.bm
} # end if(bm.dist == "NAFO")



# Finally, we can do the exact same thing with the depth stratification scheme
if(bm.dist == "Depth")
{
# Here are some effect sizes for this, should be on par with the effect size of the GRMF with the default settings
depth.strata$effect <- c(-2.50,  #<25
                         -3.00,  #> 250
                          2.00,  #100-125
                          1.50 , #125-150
                          0.50,  #150-175
                          0.25,  #175-200
                          0.00,  #200-225
                         -1.00,  #225-250
                         -1.50,  #25-50
                         -0.50,  #50-75
                          0.50)  #75-100
# Now the depth strata and our grid 'boxes' aren't going to perfectly overlap, but we can do some gymnastics to have the strata effect due to depth because
# we did the above.... Start with the points...
depth.gmrf.pnts <- st_intersection(gmrf.pnts.sf,depth.strata)
depth.gmrf.pnts$mn.bm.dens <- mn.bm.dens
depth.gmrf.pnts$bm.dens <- exp(log(depth.gmrf.pnts$mn.bm.dens) + depth.gmrf.pnts$variable1 + depth.gmrf.pnts$effect)
# Now we can take out the stratification indicators (seems to cause grief with next step)
depth.gmrf.pnts <- depth.gmrf.pnts %>% dplyr::select(!c('Group.1','layer',"area",'strata','stations','area.surveyed','towable.area','p.area'))
# Now put these into the survey grid
gmrf.bm <- st_join(survey.grid,depth.gmrf.pnts)
#calculate the area of the cells
gmrf.bm$grid.area <- gmrf.bm %>% st_area() %>% units::set_units("km^2") %>% as.numeric()
# And have fun calculating biomasses
gmrf.bm$biomass <- gmrf.bm$grid.area*gmrf.bm$bm.dens
tmp.biomass <- sum(gmrf.bm$biomass)
# Now we adjust it back to our initial biomass
gmrf.bm$biomass <- gmrf.bm$biomass*tot.biomass/tmp.biomass
gmrf.bm$bm.dens <- gmrf.bm$biomass/gmrf.bm$grid.area
bm.res[[i]] <- gmrf.bm
} # end if(bm.dist == "Depth")                        
}

```

<!-- Here we get the biomass estimates using tows that are randomly distributed across the survey area -->

```{r rand-bm, echo=F,include=F,message=F,fig.width=8,fig.height=8}

rand.tmp <- NULL
#rand.samp <- survey.domain %>% st_sample(size = n.stations) %>% st_sf()
for(i in 1:n.sims)
{
rand.samp.dat <- st_join(rand.samp[[i]],bm.res[[i]])
# This is biomass in KG at the moment, just to make it make more sense.  Note the order of magnitude difference between the size of the stock and how many
# are caught in this survey
rand.samp.dat$sample.bm <- rlnorm(n.stations,log(catchability*rand.samp.dat$bm.dens*area.swept*1000),0.25)
 # ggplot() + geom_sf(data = rand.samp.dat, aes(fill = sample.bm,color=sample.bm)) + 
 #            geom_sf(data=survey.domain, fill = NA) +
 #            scale_fill_viridis_b() + scale_color_viridis_b()

# Now from here we use standard survey math to get a survey estimated biomass.  Steal from Stephens Simple survey code because i know that is done right :-)
# The basic calculations for this are very straightforward, the average biomass per tow along with uncertainies
bm.per.tow <- mean(rand.samp.dat$sample.bm)
se.per.tow <- sd(rand.samp.dat$sample.bm)/sqrt(n.stations)
# Our DF here is the number of stations
ci.per.tow <- bm.per.tow + (c(qt(0.05/2, n.stations), -qt(0.05/2, n.stations)) * se.per.tow)

# Alternatively, we just immediately can scale up to total area..
bm.tot <- bm.per.tow* tot.towable.area /1000
ci.tot <- ci.per.tow* tot.towable.area /1000

# And here is our 'q corrected' survey biomass estimate from the random survey.
bm.tot.q.cor <- bm.tot/catchability
ci.tot.q.cor <- ci.tot/catchability
rand.tmp[[i]] <- data.frame(biomass = c(bm.per.tow,bm.tot,bm.tot.q.cor),
                       lci = c(min(ci.per.tow),min(ci.tot),min(ci.tot.q.cor)),
                       uci = c(max(ci.per.tow),max(ci.tot),max(ci.tot.q.cor)),
                       metric = c("Per Tow","Survey","Q-Corrected"),
                       survey = rep('Random',3),
                       simulation = rep(i,3))
}

rand.res <- do.call('rbind',rand.tmp)
```

<!-- Here we get the biomass estimates using tows that are distributed across the survey area according to NAFO stratification -->

```{r nafo-bm, echo=F,include=F,message=F}

nafo.tmp <- NULL
for(i in 1:n.sims)
{
# Here's a go at it, 
nafo.samp.dat <- st_join(nafo.samp[[i]],bm.res[[i]])
# This is biomass in KG at the moment, just to make it make more sense.  Note the order of magnitude difference between the size of the stock and how many
# are caught in this survey
nafo.samp.dat$sample.bm.per.tow <- rlnorm(n.stations,log(catchability*nafo.samp.dat$bm.dens*area.swept*1000),0.25)
# Biomass density in kg/km^2
nafo.samp.dat$sample.bm.dens <- nafo.samp.dat$sample.bm.per.tow/area.swept
# Now things get much more complex as we have to get stratified estimates.
#First we get the mean for each strata, the 'id' field is strata at the moment
strata.res <- nafo.samp.dat %>% dplyr::group_by(id,area,p.area,stations,area.surveyed,towable.area) %>% 
                                dplyr::summarise(mn.ptow = mean(sample.bm.per.tow),
                                                 var.ptow = var(sample.bm.per.tow))
# Now we take these strata means and account for the area in each strata.  This is the biomass per tow
strata.res  <- strata.res %>% dplyr::mutate(mn.pt.by.pa = mn.ptow*p.area)
bm.per.tow <- sum(strata.res$mn.pt.by.pa)
# Calculate the survey standard error.  In words... The Total number of towable units in a strata * difference 
# So all the area calculations are done in 1a and 2a, then we multiply by the variance, then we have to divide by the number of stations
se.calc.1 <- ((strata.res$area * (strata.res$area-strata.res$area.surveyed))) # Total area * Total area not surveyed
se.calc.2 <- se.calc.1/sum(strata.res$area)^2 # Now we divide the above by the total area squared.  
se.calc.3 <- se.calc.2 * strata.res$var.ptow # Then we multiply that by the variance 
se.survey <- sum(se.calc.3/strata.res$stations,na.rm=T)^0.5 # then we divide this total by the number of stations, add up that number and take square root.
  
# Now we need to know the difference between the area towed and the possible area towed, needed to figure out DF later...
#ah <- (strata.res$towable.area * (strata.res$towable.area - strata.res$stations))/strata.res$stations
area.df <- (strata.res$area * (strata.res$area - strata.res$area.surveyed))/strata.res$area.surveyed
# Degrees of freedom
df <- (sum(area.df * strata.res$var.ptow, na.rm = TRUE)^2)/(sum(((area.df * strata.res$var.ptow)^2)/(strata.res$stations - 1), na.rm = TRUE))
# And the size of our CI around the mean estimate (95% CI)
ci.per.tow <- bm.per.tow + (c(qt(0.05/2, df), -qt(0.05/2, df)) * se.survey)
# Now scale up to whole bank, all these calculations are done per tow, so how many 'tows' could one do in the whole region?
bm.tot <- bm.per.tow*tot.towable.area /1000 # Note we are in kg/tow, reset to tonnes for whole bank
ci.tot <- ci.per.tow * tot.towable.area /1000 # Note we are in kg/tow, reset to tonnes for whole bank
# Q-corrected
bm.tot.q.cor <- bm.tot/catchability
ci.tot.q.cor <- ci.tot/catchability
# Put it into a dataframe
nafo.tmp[[i]] <- data.frame(biomass = c(bm.per.tow,bm.tot,bm.tot.q.cor),
                       lci = c(min(ci.per.tow),min(ci.tot),min(ci.tot.q.cor)),
                       uci = c(max(ci.per.tow),max(ci.tot),max(ci.tot.q.cor)),
                       metric = c("Per Tow","Survey","Q-Corrected"),
                       survey = rep('NAFO Strata',3),
                       simulation = rep(i,3))
}
nafo.res <- do.call('rbind',nafo.tmp)

```

<!-- Here we get the biomass estimates using tows that are distributed across the survey area according to Depth stratification -->

```{r depth-bm, echo=F,include=F,message=F}

depth.tmp <- NULL
for(i in 1:n.sims)
{
depth.samp.dat <- st_join(depth.samp[[i]],bm.res[[i]])
# This is biomass in KG at the moment, just to make it make more sense.  Note the order of magnitude difference between the size of the stock and how many
# are caught in this survey
depth.samp.dat$sample.bm.per.tow <- rlnorm(n.stations,log(catchability*depth.samp.dat$bm.dens*area.swept*1000),0.25)
# Biomass density in kg/km^2
depth.samp.dat$sample.bm.dens <- depth.samp.dat$sample.bm.per.tow/area.swept
# Now things get much more complex as we have to get stratified estimates.
#First we get the mean for each strata, the 'id' field is strata at the moment
depth.strata.res <- depth.samp.dat %>% dplyr::group_by(strata,area,p.area,stations,area.surveyed,towable.area) %>% 
                                dplyr::summarise(mn.ptow = mean(sample.bm.per.tow),
                                                 var.ptow = var(sample.bm.per.tow))
# Now we take these strata means and account for the area in each strata.  This is the biomass per tow
depth.strata.res  <- depth.strata.res %>% dplyr::mutate(mn.pt.by.pa = mn.ptow*p.area)
bm.per.tow <- sum(depth.strata.res$mn.pt.by.pa)
# Calculate the survey standard error.  In words... The Total number of towable units in a strata * difference 
  # between Total number of towable units and the actual number of tows in the strata divided by square of total number 
  # of towable units on the bank, multiply all this by variance in each strata, and divided everything by 
  # the number of tows in each strata
  # Finally add this up across all strata and take the square root.  In math this is...
  # (sum((Nh_i * (Nh_i-nh_i)/sum(Nh_i)^2) * var_i) / nh_i)^0.5

# I'm going to try this in terms of swept area and towable units cause I think they'll land in the same place right
# calc.1 <- ((strata.res$towable.area * (strata.res$towable.area-strata.res$stations))) # Total
# calc.2 <- calc.1/sum(strata.res$towable.area)^2
# calc.3 <- calc.2 * strata.res$var
# calc.4 <- sum(calc.3/strata.res$stations)^0.5
# se.survey

# So all the area calculations are done in 1a and 2a, then we multiply by the variance, then we have to divide by the number of stations
se.calc.1 <- ((depth.strata.res$area * (depth.strata.res$area-depth.strata.res$area.surveyed))) # Total area * Total area not surveyed
se.calc.2 <- se.calc.1/sum(depth.strata.res$area)^2 # Now we divide the above by the total area squared.  
se.calc.3 <- se.calc.2 * depth.strata.res$var.ptow # Then we multiply that by the variance 
se.survey <- sum(se.calc.3/depth.strata.res$stations,na.rm=T)^0.5 # then we divide this total by the number of stations, add up that number and take square root.
  
# Now we need to know the difference between the area towed and the possible area towed, needed to figure out DF later...
#ah <- (strata.res$towable.area * (strata.res$towable.area - strata.res$stations))/strata.res$stations
area.df <- (depth.strata.res$area * (depth.strata.res$area - depth.strata.res$area.surveyed))/depth.strata.res$area.surveyed
# Degrees of freedom
df <- (sum(area.df * depth.strata.res$var.ptow, na.rm = TRUE)^2)/(sum(((area.df * depth.strata.res$var.ptow)^2)/(depth.strata.res$stations - 1), na.rm = TRUE))
# And the size of our CI around the mean estimate (95% CI)
ci.per.tow <- bm.per.tow + (c(qt(0.05/2, df), -qt(0.05/2, df)) * se.survey)
# Now scale up to whole bank, all these calculations are done per tow, so how many 'tows' could one do in the whole region?
bm.tot <- bm.per.tow*tot.towable.area /1000 # Note we are in kg/tow, reset to tonnes for whole bank
ci.tot <- ci.per.tow * tot.towable.area /1000 # Note we are in kg/tow, reset to tonnes for whole bank
# Q-corrected
bm.tot.q.cor <- bm.tot/catchability
ci.tot.q.cor <- ci.tot/catchability
# Put it into a dataframe
depth.tmp[[i]] <- data.frame(biomass = c(bm.per.tow,bm.tot,bm.tot.q.cor),
                       lci = c(min(ci.per.tow),min(ci.tot),min(ci.tot.q.cor)),
                       uci = c(max(ci.per.tow),max(ci.tot),max(ci.tot.q.cor)),
                       metric = c("Per Tow","Survey","Q-Corrected"),
                       survey = rep('Depth Strata',3),
                       simulation = rep(i,3))
} # end for(i in 1:n.sims)
depth.res <- do.call('rbind',depth.tmp)

```


<!-- This makes plot objects that we want for plotting the biomass distribution with the survey coverage for each of the stratification schemes -->

```{r bm-dist,echo=F, warning =F, message=F,include =F}

p.rand <- NULL
p.nafo <- NULL
p.depth <- NULL

p.rand[[1]] <- base.no.bathy + geom_sf(data=bm.res[[1]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = rand.samp[[1]],color='white') + theme_map() +theme(legend.position = "none")+
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
# Now the nafo coverage
p.nafo[[1]] <- base.no.bathy + geom_sf(data=bm.res[[1]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = nafo.samp[[1]],color='white') + theme_map() +theme(legend.position = "none")+
                            geom_sf(data = nafo.strata, fill = NA) +                          
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)

# Now the depth coverage
p.depth[[1]] <- base.no.bathy + geom_sf(data=bm.res[[1]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = depth.samp[[1]],color='white') + theme_map() + theme(legend.position = "none")+
                            geom_sf(data = depth.strata, fill = NA) +                          
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)

if(n.sims >1)
{
  # Now grab the same info for the second simulation
  p.rand[[2]] <- base.no.bathy + geom_sf(data=bm.res[[2]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = rand.samp[[2]],color='white') + theme_map() +theme(legend.position = "none")+
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # Now the nafo coverage
  p.nafo[[2]] <- base.no.bathy + geom_sf(data=bm.res[[2]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = nafo.samp[[2]],color='white') + theme_map() +theme(legend.position = "none")+
                            geom_sf(data = nafo.strata, fill = NA) +
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # Now the depth coverage
  p.depth[[2]] <- base.no.bathy + geom_sf(data=bm.res[[2]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = depth.samp[[2]],color='white') + theme_map() +theme(legend.position = "none")+
                            geom_sf(data = depth.strata, fill = NA) +  
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # If we have more than 2 simulations we also grab the last simulation and show that one too.
  

  if(n.sims >2)
  {
  # Now grab the same info for the second simulation
  p.rand[[3]] <- base.no.bathy + geom_sf(data=bm.res[[3]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = rand.samp[[3]],color='white') + theme_map() +theme(legend.position = "none")+
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # Now the nafo coverage
  p.nafo[[3]] <- base.no.bathy + geom_sf(data=bm.res[[3]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = nafo.samp[[3]],color='white') + theme_map() +theme(legend.position = "none")+
                            geom_sf(data = nafo.strata, fill = NA) +
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # Now the depth coverage
  p.depth[[3]] <- base.no.bathy + geom_sf(data=bm.res[[3]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = depth.samp[[3]],color='white') + theme_map() +theme(legend.position = "none")+
                            geom_sf(data = depth.strata, fill = NA) +  
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # If we have more than 2 simulations we also grab the last simulation and show that one too.
  }#end  if(n.sims ==3)
    
  
  if(n.sims >3)
  {
  # Now grab the same info for the final simulation
  p.rand[[4]] <- base.no.bathy + geom_sf(data=bm.res[[n.sims]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = rand.samp[[n.sims]],color='white') + theme_map() + theme(legend.position = "none")+
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # Now the nafo coverage
  p.nafo[[4]] <- base.no.bathy + geom_sf(data=bm.res[[n.sims]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = nafo.samp[[n.sims]],color='white') + theme_map() + theme(legend.position = "none")+
                            geom_sf(data = nafo.strata, fill = NA) +
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  # Now the depth coverage
  p.depth[[4]] <- base.no.bathy + geom_sf(data=bm.res[[n.sims]], aes(fill = bm.dens,color=bm.dens)) + 
                            geom_sf(data = depth.samp[[n.sims]],color='white') + theme_map() +theme(legend.position = "none")+
                            geom_sf(data = depth.strata, fill = NA,) +  
                            scale_color_viridis_c(option = 'C',direction =1) + scale_fill_viridis_c(option = 'C',direction =1)
  } #end  if(n.sims > 3)

}# ENd if(n.sims > 1)
  


```


## Survey Simulation

So now we can review the input data we have for our survey.  First we will look at some figures. First off, lets take a look at our survey area, included in this figure are the North Atlantic Fishery Organization (NAFO) subareas that are the basis for the NAFO stratification, and the bathymetry of the region, which is used as the basis of the depth stratification (Figure \@ref(fig:base-plt)).  


```{r base-plt,echo=F, warning =F, message=F,fig.width=8,fig.height=8, fig.cap="The assessment area for the Dusky Scalloped Shark (*Dustious maximus*) is outlined by the thick grey line. The thin black lines are the NAFO subareas in the region. The red line divides shows the division between the economic exclusive zone (EEZs) for Canada and the United States. The bathymetry in the region is also shown."}

base.plt 

```



Now we can also show the distribution of the biomass in the area.  If `r n.sims` is greater than 1 then we'll show two or three realizations from the models depending on how many simulations we ran. First we show the biomass distribution with the random survey stations overlain (Figure \@ref(fig:rand-samp-plt)).


```{r rand-samp-plt,echo=F, warning =F, message=F,fig.width=7,fig.height=10, fig.cap="Biomass distribution with the random survey stations overlain"}
# Now make our plot(s)
if(n.sims == 1) p.rand.plt <- p.rand[[1]]  + ggtitle(paste("Simulation 1"))
if(n.sims == 2) p.rand.plt <-   plot_grid(p.rand[[1]]+ ggtitle(paste("Simulations 1 and 2")),p.rand[[2]]+ ggtitle(""),ncol =1) 
if(n.sims ==3)  p.rand.plt <-   plot_grid(p.rand[[1]]+ ggtitle(paste("Simulations 1, 2 and 3")),p.rand[[2]]+ ggtitle(""),p.rand[[3]],ncol =2)
if(n.sims >3)   p.rand.plt <-   plot_grid(p.rand[[1]]+ ggtitle(paste("Simulations 1, 2, 3 and",n.sims)),p.rand[[2]]+ ggtitle(""),p.rand[[3]],p.rand[[4]],ncol =2)
p.rand.plt
```

Next we show the biomass distribution with the NAFO survey stations and NAFO strata overlain (Figure \@ref(fig:rand-samp-plt)). 

```{r nafo-samp-plt,echo=F, warning =F, message=F,fig.width=7,fig.height=10, fig.cap="Biomass distribution with the NAFO survey stations and NAFO stratification polygons overlain"}
# Now make our plot(s)
if(n.sims == 1) p.nafo.plt <- p.nafo[[1]] + ggtitle(paste("Simulation 1"))
if(n.sims == 2) p.nafo.plt <-   plot_grid(p.nafo[[1]]+ ggtitle(paste("Simulations 1 and 2")),p.nafo[[2]]+ ggtitle(""),ncol =1)
if(n.sims == 3) p.nafo.plt <-   plot_grid(p.nafo[[1]]+ ggtitle(paste("Simulations 1, 2 and 3")),p.nafo[[2]]+ ggtitle(""),p.nafo[[3]],ncol =2)
if(n.sims > 3)  p.nafo.plt <-   plot_grid(p.nafo[[1]]+ ggtitle(paste("Simulations 1, 2, 3 and",n.sims)),p.nafo[[2]]+ ggtitle(""),p.nafo[[3]],p.nafo[[4]],ncol =2)

p.nafo.plt
```

Finally, we show the biomass distribuiton with the Depth survey stations and Depth stratification overlain (Figure \@ref(fig:depth-samp-plt))

```{r depth-samp-plt,echo=F, warning =F, message=F,fig.width=7,fig.height=10, fig.cap="Biomass distribution with the Depth survey stations overlain and the Depth stratification polygons overlain."}
# Now make our plot(s)
if(n.sims == 1) p.depth.plt <- p.depth[[1]]+ ggtitle(paste("Simulation 1"))
if(n.sims == 2) p.depth.plt <-   plot_grid(p.depth[[1]]+ ggtitle(paste("Simulations 1 and 2")),p.depth[[2]] + ggtitle(""),ncol =1)
if(n.sims == 3) p.depth.plt <-   plot_grid(p.depth[[1]]+ ggtitle(paste("Simulations 1, 2 and 3")),p.depth[[2]]+ ggtitle(""),p.depth[[3]],ncol =2)
if(n.sims > 3)  p.depth.plt <-   plot_grid(p.depth[[1]]+ ggtitle(paste("Simulations 1, 2, 3 and",n.sims)),p.depth[[2]]+ ggtitle(""),p.depth[[3]],p.depth[[4]],ncol =2)
p.depth.plt
```

# Now we can compare the random survey estimates to the depth and NAFO stratified surveys.

```{r bm-plts,echo=F,message=F,fig.width=7,fig.height=8, fig.cap = "Biomass estimates from the 3 different survey sampling schemes.  When the number of simulations run = 1 this provides the mean and 95% CI from that simulation.  When the number of simulations is >1 and < 10 the mean biomass for each simulation is shown.  When the number of simulations is >=10 we show the median biomass of the simulations along with the interquartile range of the biomass from the simulations",cache=F}

res <- bind_rows(rand.res,nafo.res,depth.res)


res$metric <- factor(res$metric, levels = c("Per Tow","Survey","Q-Corrected"))
h.lines <- rep(rep(c(expected.bm.per.tow,expect.survey.bm,tot.biomass),3),n.sims)


# If we 10 or more simulations, go with the modified boxplot
if(n.sims >=10)
{
h.lines <- rep(c(expected.bm.per.tow,expect.survey.bm,tot.biomass),3)

b.quants <- res %>% group_by(survey,metric) %>% dplyr::summarise(med = quantile(biomass,probs = c(0.5)),
                                                                 q25 = quantile(biomass,probs = c(0.25)),
                                                                 q75 = quantile(biomass,probs = c(0.75)))
p.bm <- ggplot(b.quants) + geom_point(aes(y=med,x = survey)) + facet_wrap(~metric,scale='free_y',ncol = 1) +
                geom_errorbar(aes(x=survey,ymin=q25,ymax=q75),width=0) + ylab("Median Biomass (with Interquartile Range)")+ xlab("")+
                geom_hline(aes(yintercept = h.lines),color='grey',linetype='dashed') +
                theme_bw()
p.bm
}

# If we are running between 2 and 10 simulations plot the points only
if(n.sims >1 & n.sims <10)
{
p.bm <- ggplot(res) + geom_text(aes(y=biomass,x = survey,label = simulation)) + facet_wrap(~metric,scale='free_y',ncol = 1) +
                #geom_errorbar(aes(x=survey,ymin=lci,ymax=uci),width=0) + 
                geom_hline(aes(yintercept = h.lines),color='grey',linetype='dashed') +
                ylab("Mean Biomasfor each simulation ")+xlab("") + theme_bw()
p.bm
}

# If we are running 1 simulation then we just make plot with the CIs.
if(n.sims == 1)
{
p.bm <- ggplot(res) + geom_point(aes(y=biomass,x = survey),) + facet_wrap(~metric,scale='free_y',ncol = 1) +
                geom_errorbar(aes(x=survey,ymin=lci,ymax=uci),width=0) + xlab("")+  ylab("Mean (with 95% CI)")+
                geom_hline(aes(yintercept = h.lines),color='grey',linetype='dashed') +
                theme_bw()
p.bm
}

```


